---
title: "Studying the trajectories"
author: "Di Cook"
date: "July 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
library(tidyr)
library(dplyr)
library(purrr)
library(ggplot2)
```

## Function to compute trajectory

This function computes the position given the coordinates and a time point.

```{r}
traj_coords <- function(start, dir, flip, c0, c1, c2, c3, tm=0) {
  tm <- tm + start
  if (dir != "z") {
    c0 <- flip * c0
    c1 <- flip * c1
    c2 <- flip * c2
    c3 <- flip * c3
  }
  pos <- c0 + c1*tm + c2*tm^2 + c3*tm^3
  return(pos)
}
```

# Load and process data

Read in the data (needs to be in the same directory), and create additional variables of interest. 

```{r}
load("atp_serves.Rdata")
arc1 <- 10
arc3 <- 0
atp_serves <- atp_serves %>% mutate(duration = arc1 * duration.arc1 + arc3 * duration.arc3,
                          flip = sign(raw.x0.1), 
                          tserve = ifelse(abs(center.y)>2, "no", "yes"))
```

The data should be in long form, so that the trajectory function can be applied independently to each x, y, z direction.

```{r}
atp_serves_long <- atp_serves %>%
  select(serveid, start.1, duration.arc1, flip, raw.x0.1:raw.z3.1) %>%
  gather(coef, value, -serveid, -start.1, -duration.arc1, -flip) %>%
  #gather(coef, value, raw.x0.1:raw.z3.1) %>%
  separate(coef, c("junk1", "coef", "junk2"), sep="\\.") %>%
  select(-junk1, -junk2) %>%
  mutate(dir=substr(coef, 1, 1), coef=substr(coef, 2, 2)) %>%
  spread(coef, value) %>%
  rename("start" = `start.1`, "duration"=`duration.arc1`, "c0"=`0`,
             "c1"=`1`, "c2"=`2`, "c3"=`3`)

```

Use purrr to apply the `traj_coords` function to each direction of each serve, for a sequence of time points.

```{r}
arc1points <- seq(0, 1, length.out = arc1) %>%
  map_df(function(x) mutate(rowwise(atp_serves_long), time=start + duration * x,
     pos=traj_coords(start, dir, flip, c0, c1, c2, c3,
     tm = duration * x))) %>%
  select(serveid, pos, time, flip, dir) %>%
  arrange(serveid, pos, time) %>% 
  spread(dir, pos)
```

Merge with original data - could have kept it all together, but it feels better to just operate on the coefficients subset.

```{r}
all <- merge(arc1points, atp_serves, by="serveid")
```

# Plotting the data

Set up a theme for displaying the court, and the court marks.

```{r}
# Court background theme
theme_court <- theme_bw()
theme_court$line <- element_blank()
theme_court$axis.text <- element_blank()
theme_court$axis.title <- element_blank()

# Court dimensions data
court_trace <- data.frame(x = c(-11.89, -11.89, 0, 0, 0, 0, 0, 0, 
                               11.89, 11.89, -11.89, -11.89, 11.89, 
                               11.89, -11.89, -6.4, -6.4, 6.4, 6.4, 
                               6.4, -6.4),
                         y = c(5.49, -5.49, -5.49, -6.5, -6.5, 6.5, 
                               6.5, -5.49, -5.49, 5.49, 5.49, 4.115, 
                               4.115, -4.115, -4.115, -4.115, 4.115, 
                               4.115, -4.115, 0, 0),
                         z = c(0, 0, 0, 0, 1.09, 1.09, 0, 0, 0, 0, 0, 
                               0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
```

Put it together

```{r}
# Plot x-y
ggplot() +
  geom_path(data = court_trace, aes(x = x, y = y),
            color = 'grey90', size = 0.5) +
  geom_segment(data = court_trace, aes(x= 0, xend= 0, y= -6.5, yend= 6.5),
               size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(data=filter(all, scorer!=0), aes(x=x, y=y, group=serveid,
        colour=factor(scorer)), alpha=0.1) +
  facet_grid(serve_classification~side) +
  coord_equal() +
  theme_court +
  theme(legend.position="bottom")
ggplot(data=filter(all, scorer!=0)) +
  geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0),
               color='grey90', size=0.5) +
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='grey90') +
  geom_line(data=filter(all, scorer!=0), aes(x=x, y=z, group=serveid,
                          colour=factor(scorer)), alpha=0.1) +
  facet_grid(serve_classification~side) +
  theme_court +
  theme(legend.position="bottom")
```

# Summarise traces

```{r}
sub <- all %>% select(serveid, time, x, y, z, server,
                      receiver, scorer, side, tserve, serve_classification) %>%
  arrange(serveid, time)
sub$timeindx <- rep(1:arc1, 2000)
sub_summary <- sub %>% group_by(side, tserve, serve_classification,
                                scorer, timeindx) %>%
  summarise(xm=mean(x, trim=0.1, na.rm=T),
            ym=mean(y, trim=0.1, na.rm=T),
            zm=mean(z, trim=0.1, na.rm=T),
            xs=sd(x, na.rm=T), ys=sd(y, na.rm=T), zs=sd(z, na.rm=T))
```

and plot the average

```{r}
ggplot(data=filter(sub_summary, scorer != 0)) +
  geom_path(data = court_trace, aes(x = x, y = y),
            color = 'grey90', size = 0.5) +
  geom_segment(data = court_trace, aes(x= 0, xend= 0, y= -6.5, yend= 6.5),
               size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(data=filter(sub_summary, scorer != 0), aes(x=xm, y=ym,
                                                       group=interaction(tserve, scorer),
                                             colour=factor(scorer))) +
  facet_grid(serve_classification~side) +
  coord_equal() +
  theme_court +
  theme(legend.position="bottom")

ggplot(data=filter(sub_summary, scorer != 0)) +
  geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0),
               color='grey90', size=0.5) +
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='grey90') +
  geom_line(data=filter(sub_summary, scorer != 0),
            aes(x=xm, y=zm, group=interaction(tserve, scorer),
                colour=factor(scorer))) +
  facet_grid(serve_classification~side) +
  theme_court +
  theme(legend.position="bottom")
```

# Functionise the code better

```{r}
generate_traj_points <- function(data, npts = 10, both_arcs = TRUE, ...) {
  
  # Handle arguments for extra variables
  dots <- c("serveid", sapply(substitute(list(...))[-1], deparse))
  extravars <- match(dots[dots%in%colnames(data)], colnames(data))

  # Longify data
  data_long <- data %>%
  select(serveid, start.1, duration.arc1, flip, raw.x0.1:raw.z3.1) %>%
  gather(coef, value, -serveid, -start.1, -duration.arc1, -flip) %>%
  #gather(coef, value, raw.x0.1:raw.z3.1) %>%
  separate(coef, c("junk1", "coef", "junk2"), sep="\\.") %>%
  select(-junk1, -junk2) %>%
  mutate(dir=substr(coef, 1, 1), coef=substr(coef, 2, 2)) %>%
  spread(coef, value) %>%
  rename("start" = `start.1`, "duration"=`duration.arc1`, "c0"=`0`,
             "c1"=`1`, "c2"=`2`, "c3"=`3`) 
  
  # Compute points for arc 1
  arcpoints <- seq(0, 1, length.out = npts) %>%
    map_df(function(x) mutate(rowwise(data_long), time=start + duration * x,
      pos=traj_coords(start, dir, flip, c0, c1, c2, c3,
      tm = duration * x))) %>%
    select(serveid, pos, time, flip, dir) %>%
    arrange(serveid, pos, time) %>% 
    spread(dir, pos)
  
  # Now second arc
  if (both_arcs) {
    data_long <- data %>% select(extravars, start.3, duration.arc3, flip,
                                       raw.x0.3:raw.z3.3) %>%
      gather(coef, value, -(1:length(extravars)), -start.3, -duration.arc3, -flip) %>%
      separate(coef, c("junk1", "coef", "junk2"), sep="\\.") %>%
      select(-junk1, -junk2) %>%
      mutate(dir=substr(coef, 1, 1), coef=substr(coef, 2, 2)) %>%
      spread(coef, value) %>%
      rename("start" = `start.3`, "duration"=`duration.arc3`, "c0"=`0`, "c1"=`1`,
             "c2"=`2`, "c3"=`3`)
    
    arc3points <- seq(0, 1, length.out=npts) %>%
      map_df(function(x) mutate(rowwise(data_long), time = start + duration * x,
                                pos = traj_coords(start, dir, flip, c0, c1, c2, c3,
                                                tm=duration * x))) %>%
      select(serveid, pos, time, flip, dir) %>%
      arrange(serveid, pos, time) %>% 
    spread(dir, pos)
    arcpoints <- rbind(arcpoints, arc3points)
  }
  return(arcpoints)
}
```

Test it

```{r}
arc_points <- generate_traj_points(atp_serves)
all <- merge(arc_points, atp_serves, by="serveid")
```

Now plot it

```{r}
# Plot x-y
ggplot() +
  geom_path(data = court_trace, aes(x = x, y = y),
            color = 'grey90', size = 0.5) +
  geom_segment(data = court_trace, aes(x= 0, xend= 0, y= -6.5, yend= 6.5),
               size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(data=filter(all, scorer!=0), aes(x=x, y=y, group=serveid,
        colour=factor(scorer)), alpha=0.1) +
  facet_grid(serve_classification~side) +
  coord_equal() +
  theme_court +
  theme(legend.position="bottom")
ggplot(data=filter(all, scorer!=0)) +
  geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0),
               color='grey90', size=0.5) +
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='grey90') +
  geom_line(data=filter(all, scorer!=0), aes(x=x, y=z, group=serveid,
                          colour=factor(scorer)), alpha=0.1) +
  facet_grid(serve_classification~side) +
  theme_court +
  theme(legend.position="bottom")
```

Plot summaries by player

```{r}
# by player, too
sub <- all %>% select(serveid, time, x, y, z, server,
                      receiver, scorer, side, tserve, serve_classification) %>%
  arrange(serveid, time)
sub$timeindx <- rep(1:arc1, 2000)
sub_summary_plyr <- sub %>% group_by(side, tserve, serve_classification,
                                     scorer, timeindx, server) %>%
  summarise(xm=mean(x, trim=0.1, na.rm=T),
            ym=mean(y, trim=0.1, na.rm=T),
            zm=mean(z, trim=0.1, na.rm=T),
            xs=sd(x, na.rm=T), ys=sd(y, na.rm=T), zs=sd(z, na.rm=T))

ggplot(data=filter(sub_summary_plyr, scorer != 0 &
                     server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI"))) +
  geom_path(data = court_trace, aes(x = x, y = y),
            color = 'grey90', size = 0.5) +
  geom_segment(data = court_trace, aes(x= 0, xend= 0, y= -6.5, yend= 6.5),
               size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(data=filter(sub_summary_plyr, scorer != 0 &
                          server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI")),
            aes(x=xm, y=ym,
                                                       group=interaction(tserve, scorer),
                                                       colour=factor(scorer))) +
  facet_grid(server~side + serve_classification) +
  coord_equal() +
  theme_court +
  theme(legend.position="bottom")

ggplot(data=filter(sub_summary_plyr, scorer != 0 &
                     server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI"))) +
  geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0),
               color='grey90', size=0.5) +
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='grey90') +
  geom_line(data=filter(sub_summary_plyr, scorer != 0 &
                          server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI")),
            aes(x=xm, y=zm, group=interaction(tserve, scorer),
                colour=factor(scorer))) +
  facet_grid(server~side + serve_classification) +
  theme_court +
  theme(legend.position="bottom")

ggplot(data=filter(sub_summary_plyr, scorer != 0 &
                     server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI"))) +
  geom_path(data = court_trace, aes(x = x, y = y),
            color = 'grey90', size = 0.5) +
  geom_segment(data = court_trace, aes(x= 0, xend= 0, y= -6.5, yend= 6.5),
               size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(data=filter(sub_summary_plyr, scorer != 0 &
                          server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI")),
            aes(x=xm, y=ym,
                                                       group=interaction(server, tserve),
                                                       colour=server)) +
  facet_grid(side + scorer ~ serve_classification) +
  coord_equal() +
  theme_court +
  theme(legend.position="bottom")

ggplot(data=filter(sub_summary_plyr, scorer != 0 &
                     server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI"))) +
  geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0),
               color='grey90', size=0.5) +
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='grey90') +
  geom_line(data=filter(sub_summary_plyr, scorer != 0 &
                          server %in% c("DJOKOVIC", "FEDERER", "RAONIC", "NISHIKORI")),
            aes(x=xm, y=zm, group=interaction(tserve, server),
                colour=server)) +
  facet_grid(side + scorer ~ serve_classification) +
  theme_court +
  theme(legend.position="bottom")
```
