---
author:
  - name: Alwin Wang
    affiliation: Monash University
    address: >
      First line
      Second line
    email: awan39@student.monash.edu
    url: http://rstudio.com
title:
  formatted: "Visualising Serve Trajectories in High-Performance Tennis with R"
  # If you use tex in the formatted title, also supply version without
  plain:     "Visualising Serve Trajectories in High-Performance Tennis with R"
  # For running headers, if needed
  short:     "Visualising Serves"
abstract: >
  This paper investigates methods to effectively visualise key characteristics of serves using trajectory of elite tennis athletes provided by Tennis Australia. The key characteristics identified were the position, velocity and spin of a ball as well as the location of single and multiple serve clusters. For the visuals presented in the paper, a sample of 2000 serves from the 2016 Australian Open from thirty three servers was used.  
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}", REMEMBER TO ADD TITLES TO YOUR GRAPHS]
  plain:     [keywords, not capitalized, Java, REMEMBER TO ADD TITLES TO YOUR GRAPHS]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

# Introduction

This template demonstrates some of the basic latex you'll need to know to create a JSS article.

```{r packages, echo=FALSE, messages=FALSE, warning=FALSE, include=FALSE}
#--- Packages Required
require(MASS)
require(cowplot)
require(lazyeval)
require(tidyr)
require(dplyr)
require(purrr)
require(broom)
require(ggplot2)
require(plotly)
require(hextri)
require(hexbin)
require(RColorBrewer)
require(aplpack)
```
```{r scripts, echo = FALSE, messages=FALSE, warning=FALSE}
#--- Load functions from helper scripts
source("Helper Load Data.R")
source("Helper Plotting Factors.R")
source("Helper Standardise Coefficients.R")
source("Helper Plotting Values.R")
source("Helper Spin Coefficients.R")
source("Helper Basic Plots.R")
source("Helper Contour Limits.R")
source("Helper Heat Map.R")
```

```{r generatedata, echo = FALSE, messages=FALSE, warning=FALSE, cache=TRUE}
#--- Generate the data for plotting
data <- PlottingFactors(atp_serves)
coef.df <- StandardiseCoefficients(data,matchid,server,speedkmph,speed_class,serve_num,serve_classname,side,scorername,
                                   start.x, start.y, start.z, center.x, center.y)
values <- PlottingValues(coef.df,tstep = 0.04)
resultstidy <-  SpinModel(values)
```

```{r tidydata, echo = FALSE, message=FALSE, warning=FALSE, cache=TRUE}
serve_outliers <- data.frame(serveid = c(
    "1_09_02_1_191947.trj", "2_04_01_1_172646.trj", "2_06_02_1_214102.trj",
    "3_01_03_2_181341.trj", "3_04_03_1_205313.trj", "3_06_08_2_174140.trj",
    "4_02_02_1_180400.trj", "4_03_05_1_190717.trj", "4_05_01_1_213913.trj",
    "1_02_01_1_161204.trj"))
# Filter out outliers
plot_sample <- resultstidy %>% 
    filter(!(serveid %in% serve_outliers$serveid))
# Data per serve
plot_perserve <- plot_sample  %>% 
    select(-(t:a), -(ax_p:az_p)) %>%
    group_by(serveid, arc) %>%
    mutate(Cd_ave = mean(Cd), Cl_ave = mean(Cl)) %>%
    distinct(.keep_all = TRUE) %>%
    gather(key, value, -(1:(match("arc", colnames(.))))) %>%
    unite(key, c(key,arc), sep="_arc") %>%
    spread(key, value) %>%
    distinct(.keep_all = TRUE)
#--- Select players with multiple games
MultipleGames <- function(min, perserve=FALSE, server=FALSE) {
    multiple_games <- atp_serves %>% 
        select(server, matchid) %>% distinct() %>% 
        count(server, sort=TRUE) %>% filter(n >= min)
    if (perserve==FALSE) {
    plot_sample %>%
        filter(server %in% multiple_games$server)
    }
    else {
      if (server==FALSE) {
      plot_perserve %>%
        filter(server %in% multiple_games$server)
      }
      else {
        as.data.frame(multiple_games %>% select(server))
      }
    }
}
```

```{r test}
head(values$t, n = 20)
head(plot_sample$t,n = 20)
```


# Shape
## Position
asdfaslkfj asfldkjasd f
```{r all_topdown, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "All serves as overlayed lines"}
plot_gg <- plot_sample
# Top Down
court_topdown + geom_path(aes(x=px,y=py, group=serveid, colour = server), data=plot_gg, alpha = 1) +  
  theme(legend.key.size = unit(0.3, "cm")) 
```

```{r all_topdown_hexbin, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Frequency of serves in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- plot_sample 
court_topdown + 
    stat_binhex(data = plot_gg, aes(x=px, y=py), bins = 20, 
                colour = "white", alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(4, "Blues"), 
                         guide = "colourbar",
                         name = "Frequeny")
```

```{r all_topdown_height, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Average height of serves in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- plot_sample
court_topdown + 
    stat_summary_hex(data = plot_gg, aes(x=px, y=py, z=pz), bins = 30, 
                fun = function(z) mean(z),
                colour = "white", alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(8, "PuRd"), 
                         guide = "colourbar",
                         name = "Ave. \nHeight") 
```


```{r all_topdown_speed, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Average impact speed in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- plot_sample
court_topdown + 
    stat_summary_hex(data = plot_gg, aes(x=px, y=py, z=speedkmph), bins = 30, 
                fun = function(z) mean(z),
                colour = "white", alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlOrBr"), 
                         guide = "colourbar",
                         name = "Average \nimpact \nspeed \n(km/h)")
```

```{r all_topdown_vel, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Average true velocity in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- plot_sample
court_topdown + 
    stat_summary_hex(data = plot_gg, aes(x=px, y=py, z=v), bins = 30, 
                fun = function(z) mean(z),
                colour = "white", alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(8, "Oranges"), 
                         guide = "colourbar",
                         name = "Average \nvelocity")
```


```{r all_topdown_side, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Scorer in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- filter(plot_sample, scorername != "No Winner")
court_topdown + 
    stat_summary_hex(data = plot_gg, aes(x=px, y=py, z=scorername), bins = 30, 
                fun = function(x) {
                   tab <- table(x)
                   names(tab)[which.max(tab)]},
                colour = "white", alpha = 0.25) +
    scale_fill_discrete(name = "Most \nOccurring")
```

```{r all_topdown_classname, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Serve type in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- filter(plot_sample, serve_classname != "Fault")
court_topdown + 
    stat_summary_hex(data = plot_gg, aes(x=px, y=py, z=serve_classname), bins = 30, 
                fun = function(x) {
                   tab <- table(x)
                   names(tab)[which.max(tab)]},
                colour = "white", alpha = 0.25) +
    scale_fill_discrete(name = "Most \nOccurring")
```

```{r all_topdown_hextriclass, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Serve type using Hextri"}
plot_gg <- filter(plot_sample, serve_classname != "Fault")
hextri(x=plot_gg$px, y=plot_gg$py,
       class=plot_gg$serve_classname, 
       colours=c("blue","orange","green", "red"), nbins = 30, border=TRUE,style="size")
    lines(x=court_trace$x, y=court_trace$y, lwd = 2)
    legend(2,10,fill=c("blue","green","orange"),
       legend=c("Ace", "Not Returned", "Returned"),bty="n")
# I could also look at a bubblechart - size is the 3rd variable
```

## Landing
```{r all_topdown_landsplitscorer, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Split landing position coloured by scorer"}
#--- Break into before bounce and after bounce
plot_arc1 <- plot_sample %>% 
    filter(arc == 1) %>% group_by(serveid) %>%
    top_n(4,-t) %>% top_n(3,t)
plot_arc3 <- plot_sample %>% 
    filter(arc == 3) %>% group_by(serveid) %>%
    top_n(4,-t) %>% top_n(2,t)
plot_gg <- plot_sample

#--- Plot
court_topdown + 
    geom_point(data = filter(plot_gg, serve_classname != "Fault"), aes(x=center.x, y=center.y, colour = scorername), alpha=0.5, shape=1) +
    geom_path(data = filter(plot_arc1, serve_classname != "Fault"), aes(x=px,y=py, group=serveid, colour = scorername), alpha = 0.5) +
    geom_path(data = filter(plot_arc3, serve_classname != "Fault"), aes(x=px,y=py, group=serveid, colour = scorername), alpha = 0.5) + 
  guides(colour = guide_legend(override.aes = list(size = 10))) +
  theme(legend.key=element_rect(fill=NA))
```

```{r all_topdown_landsplitnum, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Split landing position coloured by serve number"}
#--- Break into before bounce and after bounce
plot_arc1 <- plot_sample %>% 
    filter(arc == 1) %>% group_by(serveid) %>%
    top_n(4,-t) %>% top_n(3,t)
plot_arc3 <- plot_sample %>% 
    filter(arc == 3) %>% group_by(serveid) %>%
    top_n(4,-t) %>% top_n(2,t)
plot_gg <- plot_sample

#--- Plot
court_topdown + 
    geom_point(data = filter(plot_gg, serve_classname != "Fault"), aes(x=center.x, y=center.y, colour = serve_num), alpha=0.5, shape=1) +
    geom_path(data = filter(plot_arc1, serve_classname != "Fault"), aes(x=px,y=py, group=serveid, colour = serve_num), alpha = 0.5) +
    geom_path(data = filter(plot_arc3, serve_classname != "Fault"), aes(x=px,y=py, group=serveid, colour = serve_num), alpha = 0.5) +
   guides(colour = guide_legend(override.aes = list(size = 10))) +
   theme(legend.key=element_rect(fill=NA))
```

```{r all_landingcontoursgauss_vs_kde2d, echo=TRUE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Countour plot of all serves using Gauss density2D and KDE2d with limits"}
#--- Plotting uing geom_density2d from ggplot2
plot_gg <- plot_perserve %>% filter(serve_classname != "Fault")
plot1 <- court_service + 
    geom_density2d(data = plot_gg, aes(x=center.x, y=center.y))
#--- Plotting using Two Dimensional Kernel Density Estimation from MASS 
#    (axis-aligned bivariate normal kernel on a square grid)
plot_Ad <- plot_perserve %>% filter(side == "Ad")
densAd <- kde2d(plot_Ad$center.x, plot_Ad$center.y, 
              lims = c(-6.4, 0, 0, 4.115))
plot_Ad <- data.frame(expand.grid(center.x = densAd$x, center.y = densAd$y),
                     z = as.vector(densAd$z)) %>%
              mutate(side = "Ad")
plot_Du <- plot_perserve %>% filter(side == "Deuce")
densDu <- kde2d(plot_Du$center.x, plot_Du$center.y, 
              lims = c(-6.4, 0, -4.115, 0))
plot_Du <- data.frame(expand.grid(center.x = densDu$x, center.y = densDu$y),
                     z = as.vector(densDu$z)) %>%
              mutate(side = "Deuce")
plot2 <- court_service + 
  geom_contour(aes(x=center.x, y=center.y,z=z), data = plot_Ad) +
  geom_contour(aes(x=center.x, y=center.y,z=z), data = plot_Du)
#--- plot the results side by side
plot_grid(plot1, plot2, align='h', labels=c('Gauss', 'KDE2D'))
```

```{r landingcountours function, echo=TRUE, cache=TRUE}
```

```{r landing_contours4gamesguass, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 288, fig.align= "center", fig.cap = "Landing positiions based on serve and serve number using guass density2d"}
plot_gg <- MultipleGames(4,perserve = TRUE) %>% filter(serve_classname != "Fault")
court_service + 
    geom_density2d(data = plot_gg, aes(x=center.x, y=center.y)) +
    facet_grid(serve_num~server)
```


```{r landing_contours4gameskde, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 288, fig.align= "center", fig.cap = "Landing positions based on serve and serve number using kde2d with limits"}
ContourLimits(plot_perserve, argument1 = "serve_num", min_games = 4, plot = TRUE)
```


```{r landing_contours3gameskde, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 288, fig.align= "center", fig.cap = "Landing positions based on serve and serve number using kde2d with limits"}
plotall <-  ContourLimits(plot_perserve, argument1 = "serve_num", min_games = 3, plot = FALSE)
plotall + facet_wrap(serve_num ~ server)
```

```{r landing_contoursinteresting, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 9, dpi = 288, fig.align= "center", fig.cap = "Landing positions based on serve and serve number using kde2d with limits"}
# atp_serves %>% ungroup %>% select(loser_name,loser_rank) %>% distinct() %>% arrange(loser_rank)
# servers <- atp_serves %>% ungroup %>% select(server) %>% distinct()
servers <- data.frame(servers = c("DJOKOVIC", "MURRAY", "FEDERER", "WAWRINKA"))
plot1 <- ContourLimits(plot_perserve, "serve_num", servers, plot = TRUE)
servers <- data.frame(servers = c("SELA", "MILLMAN", "HERBERT", "ROBERT"))
plot2 <- ContourLimits(plot_perserve, "serve_num", servers, plot = TRUE)

plot_grid(plot1, plot2, align='h', 
          labels=c('Top', 'Bottom'))
          # labels=c('Top 4 Players \n(Rank: 1, 3, 2, 4)', 
          #          'Bottom 4 Players \n(Rank: 167, 95, 225, 87)'))
```

## Speed
Speed distribution
hexplot with speed average 
```{r speed_all, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of all non-fault serves categorised by the serve number and winner"}
plot_gg <- plot_perserve %>% filter(scorername != "No Winner")
ggplot(data = plot_gg, aes(speedkmph, fill=scorername, color = scorername)) + 
    geom_density(alpha = 0.2) + facet_grid(serve_num ~ scorername) +
    theme_light()
```

```{r speed_games4num, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_multiple_games <- MultipleGames(4)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server, scorername != "No Winner")
ggplot(data = plot_gg, aes(speedkmph, fill=scorername)) + 
    geom_density(alpha = 0.2) + facet_wrap(~serve_num)
```

```{r speed_games4servnum, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_multiple_games <- MultipleGames(4)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server, scorername != "No Winner")
ggplot(data = plot_gg, aes(speedkmph)) + 
   geom_density(alpha = 0.2) + 
   # geom_density(data = select(plot_gg, -scorername), aes(speedkmph)) +
   facet_grid(serve_num ~ server, labeller = label_value, scales = "free_y")
```


```{r speed_games4servnumsco, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_multiple_games <- MultipleGames(4)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server, scorername != "No Winner")
ggplot(data = plot_gg, aes(speedkmph, fill=scorername)) + 
   geom_density(alpha = 0.2) + 
   # geom_density(data = select(plot_gg, -scorername), aes(speedkmph)) +
   facet_grid(serve_num ~ server, labeller = label_value, scales = "free_y")
```

```{r speed_games4servnumside, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_multiple_games <- MultipleGames(4)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server)
ggplot(data = plot_gg, aes(speedkmph, fill=side)) + 
   geom_density(alpha = 0.2) + 
   facet_grid(serve_num ~ server, labeller = label_value, scales = "free_y")
```

```{r speed_landinghex_side100, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Average impact speed in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- plot_sample %>% filter(serve_classname != "Fault")
court_service + 
    stat_summary_2d(data = plot_gg, aes(x=center.x, y=center.y, z=speedkmph), bins = 100, 
                fun = function(z) mean(z),alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlOrBr"), 
                         guide = "colourbar",
                         name = "Average \nimpact \nspeed \n(km/h)") +
  facet_wrap(~side)
```


```{r speed_landinghex_side10, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Average impact speed in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- plot_sample %>% filter(serve_classname != "Fault")
court_service + 
    stat_summary_2d(data = plot_gg, aes(x=center.x, y=center.y, z=speedkmph), bins = 20, 
                fun = function(z) mean(z),alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlOrBr"), 
                         guide = "colourbar",
                         name = "Average \nimpact \nspeed \n(km/h)") +
  facet_wrap(~side)
```

```{r speed_landinghex_class, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Average impact speed in hexagonal bins"}
#--- Top down using hexbins
plot_gg <- plot_sample
court_service + 
    stat_summary_hex(data = plot_gg, aes(x=center.x, y=center.y, z=speedkmph), bins = 45, 
                fun = function(z) mean(z),alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlOrBr"), 
                         guide = "colourbar",
                         name = "Average \nimpact \nspeed \n(km/h)") +
  facet_wrap(~serve_classname)
```

```{r speed_games4_landinghex_num, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Average impact speed in hexagonal bins"}
plot_multiple_games <- MultipleGames(4)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server, serve_classname != "Fault")
court_service + 
    stat_summary_hex(data = plot_gg, aes(x=center.x, y=center.y, z=speedkmph), bins = 30, 
                fun = function(z) mean(z),alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(9, "YlOrBr"), 
                         guide = "colourbar",
                         name = "Average \nimpact \nspeed \n(km/h)") +
  facet_grid(serve_num ~ server)
```

```{r speed_games4_landinghex_contour_num, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 288, fig.align= "center", fig.cap = "Landing positions based on serve and serve number using kde2d with limits"}
plotall <- ContourLimits(plot_perserve, argument1 = "serve_num", min_games = 4, plot = FALSE)
plotall +
    stat_summary_hex(data = plot_gg, aes(x=center.x, y=center.y, z=speedkmph), bins = 30, 
                fun = function(z) mean(z),alpha = 0.7) +
    scale_fill_gradientn(colours = brewer.pal(5, "RdPu"), 
                         guide = "colourbar",
                         name = "Average \nimpact \nspeed \n(km/h)") +
  facet_grid(serve_num ~ server)
```

```{r speed_games4_landingpoint_contour_num, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 288, fig.align= "center", fig.cap = "Landing positions based on serve and serve number using kde2d with limits"}
plotall <- ContourLimits(plot_perserve, argument1 = "serve_num", min_games = 4, plot = FALSE)
plot_multiple_games <- MultipleGames(4)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server, serve_classname != "Fault")
plotall +
   geom_point(data = plot_gg, aes(x=center.x, y=center.y,colour=speedkmph, alpha=0.8)) +
  facet_grid(serve_num ~ server)
```


```{r speed_allsidenumspeedz, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve
ggplot(data = plot_gg, aes(start.z, fill=speed_class)) + 
   geom_density(alpha = 0.2) + 
   facet_grid(serve_num ~ side, scales = "free_y")
```


```{r speed_allsidenumspeedx, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve
ggplot(data = plot_gg, aes(start.x, fill=speed_class)) + 
   geom_density(alpha = 0.2) + 
   facet_grid(serve_num ~ side, scales = "free_y")
```


```{r speed_allsidenumspeedy, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve
ggplot(data = plot_gg, aes(start.y, fill=speed_class)) + 
   geom_density(alpha = 0.2) + 
   facet_grid(serve_num ~ side, scales = "free")
```

```{r speed_allsidenumspeedy_novak, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve %>% filter(server == "DJOKOVIC")
ggplot(data = plot_gg, aes(start.y, fill=speed_class)) + 
   geom_density(alpha = 0.2) + 
   facet_grid(serve_num ~ side, scales = "free")
```

```{r speed_allsidenumspeedy_kyrgios, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve %>% filter(server == "KYRGIOS")
ggplot(data = plot_gg, aes(start.y, fill=speed_class)) + 
   geom_density(alpha = 0.2) + 
   facet_grid(serve_num ~ side, scales = "free")
```

```{r speed_allsidenumspeedy_wawrinka, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve %>% filter(server == "WAWRINKA")
ggplot(data = plot_gg, aes(start.y, fill=speed_class)) + 
   geom_density(alpha = 0.2) + 
   facet_grid(serve_num ~ side, scales = "free")
```

```{r bubble_allnumsidespeed, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 7, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve
HeatMap(plot_gg, argument1 = "serve_num",
        argument2 = "side", argument3 = "speedkmph",
        bins=10, plot=TRUE, progress=FALSE)
```

```{r bubble_allnumclassspeed, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 8, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
plot_gg <- plot_perserve
HeatMap(plot_gg, argument1 = "serve_num",
         argument2 = "serve_classname", argument3 = "speedkmph",
         bins=10, plot=TRUE, progress=FALSE)
```


```{r bubble_top4numspeed, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 8, dpi = 144, fig.align= "center", fig.cap = "Speed of top 4 players' non-fault serves categorised by the serve number"}
HeatMap(plot_perserve, argument1 = "serve_num", min_games = 4,
                 argument3 = "speedkmph",
                 bins=5, plot = TRUE, progress = FALSE)
```

# Spin

```{r coolarc_behind, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Example of the effect of spin"}
#--- Plot single serve
serve_coolarc <- c("5_02_02_1_190119.trj","3_01_02_1_171036.trj")
plot_gg <- plot_sample %>% filter(serveid %in% serve_coolarc)
# From Behind
court_behind + geom_path(aes(x=py,y=pz, group=serveid, colour = server), data=plot_gg, alpha = 1) + 
    geom_path(data = net_trace, aes(x = y, y = z), color = 'blue', size = 1.0)
```

```{r spin_allrpm1vrpm3, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Example of the effect of spin"}
plot_gg <- plot_perserve %>% filter(w_rpm_arc1 < 5500, w_rpm_arc3 < 7500)
ggplot(plot_gg, aes(x=w_rpm_arc1, y=w_rpm_arc3, colour=speedkmph)) + geom_point()
```


```{r spin_allrpmside, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 5, dpi = 144, fig.align= "center", fig.cap = "Example of the effect of spin"}
plot_gg <- plot_perserve %>% filter(w_rpm_arc1 < 5500)
ggplot(plot_gg, aes(w_rpm_arc1, colour=serve_num)) + 
  geom_density() + 
  facet_wrap(~side)
```

```{r spin_top4rpmspeedside, echo=FALSE, cache=TRUE, fig.width = 9, fig.height = 11, dpi = 144, fig.align= "center", fig.cap = "Example of the effect of spin"}

plot_multiple_games <- MultipleGames(4)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server,
                                    w_rpm_arc1 < 5500)
plot1 <- ggplot(plot_gg, aes(w_rpm_arc1, colour=serve_num)) + 
  geom_density() + 
  facet_grid(side ~ server, scales = "free_y")
plot2 <-ggplot(plot_gg, aes(speedkmph, colour=serve_num)) + 
  geom_density() + 
  facet_grid(side ~ server, scales = "free_y")

plot_grid(plot1, plot2, align='h', ncol=1,
          labels=c('Spin', 'Speed'))
```


## Code formatting

Don't use markdown, instead use the more precise latex commands:

* \proglang{Java}
* \pkg{plyr}
* \code{print("abc")}

# R code

Can be inserted in regular R markdown blocks.

```{r}
x <- 1:10
x
```
