---
title: "Computing Ball Locations"
author: "Stephanie Kovalchik"
date: "28 June 2016"
output: html_document
---

```{r setup, eval = FALSE}
#library(devtools)
#install_github("hadley/purrr")
```

Belos is an example of how to use `purrr`, `dplyr` and the `by_row` function to compute a set of locations in one direction of a serve arc for every fixed number of seconds from the start to the end of the recording time. In the example below, we compute the x locations for every 0.05 seconds of the first arc of the serve. These are stored as a data frame of locations and times under the column `xlocation`.

```{r locations, message=FALSE}
library(dplyr)
library(purrr)

load("~/Courses/mvparrot/data/atp_serves.RData")

get_location <- function(a, b, c, d, time) c(a, b, c, d) %*% time ^ (0:3)
get_location <- Vectorize(get_location)

locations <- function(a, b, c, d, start, end, interval = 0.05) {
  times <- seq(start, end, by = interval)
  data.frame(time = times, location = get_location(a, b, c, d, times))
}

temp <- atp_serves[1:3,] %>%
  by_row(~locations(.$raw.x0.1, .$raw.x1.1, .$raw.x2.1, .$raw.x3.1, .$start.1, .$start.1 + .$duration.arc1), .to = "xlocation")


temp$xlocation
```