---
title: "Visualising Tennis Serves"
author: "Alwin"
date: "18 July 2016"
output:
    ioslides_presentation
---

```{r packages, echo=FALSE, messages=FALSE, warning=FALSE, include=FALSE}
#--- Packages Required
require(tidyr)
require(dplyr)
require(purrr)
require(broom)
require(ggplot2)
require(plotly)
require(hextri)
require(hexbin)
require(aplpack)
```

```{r, scripts, echo = FALSE, messages=FALSE, warning=FALSE}
source("Helper Load Data.R")
source("Helper Plotting Factors.R")
source("Helper Standardise Coefficients.R")
source("Helper Plotting Values.R")
source("Helper Spin Coefficients.R")
source("Helper Basic Plots.R")
```
```{r, initial functions, echo = FALSE, messages=FALSE, warning=FALSE, cache=TRUE}
data <- PlottingFactors(atp_serves)
coef.df <- StandardiseCoefficients(data,matchid,server,speedkmph,speed_class,serve_num,serve_classname,side,scorername,
                                   start.x, start.y, start.z, center.x, center.y)
values <- PlottingValues(coef.df,tstep = 0.08)
resultstidy <-  SpinModel(values)
serve_outliers <- data.frame(serveid = c(
    "1_09_02_1_191947.trj", "2_04_01_1_172646.trj", "2_06_02_1_214102.trj",
    "3_01_03_2_181341.trj", "3_04_03_1_205313.trj", "3_06_08_2_174140.trj",
    "4_02_02_1_180400.trj", "4_03_05_1_190717.trj", "4_05_01_1_213913.trj"))
# Filter out outliers
plot_sample <- resultstidy %>% 
    filter(!(serveid %in% serve_outliers$serveid))
# Data per serve
plot_perserve <- plot_sample  %>% 
    select(-(t:a), -(ax_p:az_p)) %>%
    group_by(serveid, arc) %>%
    mutate(Cd_ave = mean(Cd), Cl_ave = mean(Cl)) %>%
    distinct(.keep_all = TRUE) %>%
    gather(key, value, -(1:(match("arc", colnames(.))))) %>%
    unite(key, c(key,arc), sep="_arc") %>%
    spread(key, value) %>%
    distinct(.keep_all = TRUE)
#--- Select players with multiple games
MultipleGames <- function(min) {
    multiple_games <- atp_serves %>% 
        select(server, matchid) %>% distinct() %>% 
        count(server, sort=TRUE) %>% filter(n >= min)
    plot_sample %>%
        filter(server %in% multiple_games$server)
}
```

## Looking at all the Data {.flexbox .vcenter}
```{r coolarc_topdown, echo=FALSE, cache=TRUE}
plot_gg <- plot_sample
# Top Down
court_topdown + geom_path(aes(x=px,y=py, group=serveid, colour = server), data=plot_gg, alpha = 1)
```

## Looking at some of the Data 
Clearly there is a lot going on in serve data!
<div class="centered">
```{r cool_behind, echo=FALSE, cache=TRUE}
#--- Plot single serve
serve_coolarc <- c("5_02_02_1_190119.trj","3_01_02_1_171036.trj")
plot_gg <- plot_sample %>% filter(serveid %in% serve_coolarc)
# From Behind
court_behind + geom_path(aes(x=py,y=pz, group=serveid, colour = server), data=plot_gg, alpha = 1) + 
    geom_path(data = net_trace, aes(x = y, y = z), color = 'blue', size = 1.0)
```
</div>

## Look at the Data
So clearly there is a lot going on in Tennis serve data!
<div class="columns-2">

> - Position
    - Time, x, y, z coordinates
> - Speed
    - Speed on impact, speed throughout the serve
> - Curvature
    - Spin on the ball, Wind effects
> - Location
    - Single serve, Serve cluster
    
</div>

## Ways to Visualise Multivariable Data 
> There are good ways ...
> And bad ways...

## Not so good way {.flexbox .vcenter}
```{r chernoff, echo=FALSE, cache=TRUE}
faces(atp_serves[1:12,3:30])
```


## Position - Frequency {.flexbox .vcenter}
```{r all_topdown_hexbin, echo=FALSE, cache=TRUE}
#--- Top down using hexbins
plot_gg <- plot_sample
court_topdown + 
    stat_binhex(data = plot_gg, aes(x=px, y=py), bins = 25, colour = "gray", alpha = 0.7) +
    scale_fill_gradientn(colours = c("yellow","orange","red"))
```

## Position - Coloured by a variable {.flexbox .vcenter}
```{r all_topdown_hextri, echo=FALSE, cache=TRUE}
plot_gg <- plot_sample
hextri(x=plot_gg$px, y=plot_gg$py,
       class=plot_gg$serve_num, 
       colours=c("blue","green"), nbins = 25, border=TRUE,style="size")
    lines(x=court_trace$x, y=court_trace$y, lwd = 2)
    legend("topright",fill=c("blue","green"),
       legend=c("First Serve", "Second Serve"),bty="n")
```

## Landing Position {.flexbox .vcenter}
```{r all_land_split, echo=FALSE, cache=TRUE}
#--- Break into before bounce and after bounce
plot_arc1 <- plot_sample %>% 
    filter(arc == 1) %>%
    group_by(serveid) %>%
    top_n(4,t) %>%
    top_n(2,-t)
plot_arc3 <- plot_sample %>% 
    filter(arc == 3) %>%
    group_by(serveid) %>%
    top_n(4,-t) %>%
    top_n(2,t)
plot_gg <- plot_sample

#--- Plot
court_topdown + 
    geom_point(data = filter(plot_gg, serve_classname != "Fault"), aes(x=center.x, y=center.y, colour = scorername), alpha=0.5) +
    geom_path(data = filter(plot_arc1, serve_classname != "Fault"), aes(x=px,y=py, group=serveid, colour = scorername), alpha = 0.5) +
    geom_path(data = filter(plot_arc3, serve_classname != "Fault"), aes(x=px,y=py, group=serveid, colour = scorername), alpha = 0.5)
```

## Landing Position {.flexbox .vcenter}
```{r landing_contours, echo=FALSE, cache=TRUE}
plot_multiple_games <- MultipleGames(5)
plot_gg <- plot_perserve %>% filter(server %in% plot_multiple_games$server)
court_service + 
    geom_density2d(data = plot_gg, aes(x=center.x, y=center.y)) +
    facet_wrap(serve_num~server)
```

## Landing Position with Overlays {.flexbox .vcenter}

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

