---
title: "Untitled"
author: "Di Cook"
date: "June 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(ggplot2)
library(dplyr)
library(purrr)
```

## Load data

```{r cars}
load("atp_serves.RData")
```

## Including Plots

Starting position from a top down view for Federer and Djokovic.
```{r fed-djo, echo=FALSE}
fed <- atp_serves %>% filter(server == "FEDERER")
ggplot(fed, aes(x=start.x, y=start.y)) + geom_point()

djok <- atp_serves %>% filter(server == "DJOKOVIC")
ggplot(djok, aes(x=start.x, y=start.y)) + geom_point()
```

# Examine the contact point of serves

Landing position from a top down view for all serves.
```{r}
ggplot(data=atp_serves, aes(x=center.x, y=center.y)) + 
  geom_point() 
ggplot(data=atp_serves, aes(x=center.x, y=center.y)) + 
  geom_point() + facet_wrap(~side)
```

# Examine winners

Contour plot of landing position. Columns of Ad and Deuce. Rows of scorer.
Overlay of two contour plots of scorer. Columns of Ad and Deuce.
```{r}
ggplot(data=filter(atp_serves, scorer != 0), aes(x=center.x, y=center.y)) + 
  geom_point() + geom_density2d() + facet_grid(scorer~side)
ggplot(data=filter(atp_serves, scorer != 0), 
       aes(x=center.x, y=center.y, colour=factor(scorer))) + 
  geom_point() + geom_density2d() + facet_wrap(~side)
```

# By server??

Two overlayed countour plots based by scorer. Columns of Ad and Deuce. Rows by server.
```{r}
keep <- atp_serves %>% select(server, matchid) %>% 
  distinct() %>% count(server, sort=TRUE) %>% filter(n>2) 
atp_serves_sub <- atp_serves %>% filter(server %in% keep$server)
ggplot(data=filter(atp_serves_sub, scorer != 0), 
       aes(x=center.x, y=center.y, colour=factor(scorer))) + 
  geom_point() + geom_density2d(alpha=0.5) + facet_grid(server~side)
```

# Aces by server

All aces showed in grey. Aces for specific servers are highlighted. Columns are Ad and Deuce. Rows are Server.
```{r}
all <- atp_serves_sub %>% filter(serve_classification == 0) %>%
  select(side, center.x, center.y)
servers <- atp_serves_sub %>% filter(serve_classification == 0) %>%
  select(side, center.x, center.y, server)
ggplot(data=all, aes(x=center.x, y=center.y)) +
  geom_point(alpha=0.2) +
  geom_point(data=servers, aes(x=center.x, y=center.y, colour=server), size=2) + 
  facet_grid(server~side)
```

# Returned by server, compared to aces

Landing position of non-fault serves with aces highlighted. Columns are Ad and Deuce. Rows are Server.
```{r}
servers <- atp_serves_sub %>% filter(serve_classification < 2) %>%
  select(side, center.x, center.y, server, serve_classification)
ggplot(data=servers, aes(x=center.x, y=center.y)) +
  geom_point(data=servers, 
             aes(colour=factor(serve_classification)), alpha=0.5, size=2) + 
  facet_grid(server~side) + theme(legend.position="none")
```

# Returned by server, and receiver

Playing around with axis
```{r}
servers <- atp_serves_sub %>% filter(serve_classification < 2) %>%
  select(side, center.x, center.y, server, serve_classification)
ggplot(data=servers, aes(x=center.y, y=center.x)) +
  geom_point(data=servers, 
             aes(colour=factor(serve_classification)), alpha=0.5, size=2) + 
  facet_grid(side~server) + theme(legend.position="none")

servers <- atp_serves_sub %>% filter(serve_classification < 2) %>%
  select(side, center.x, center.y, server, serve_classification)
ggplot(data=servers, aes(x=center.y, y=center.x)) +
  geom_point(data=servers, 
             aes(colour=factor(serve_classification)), alpha=0.5, size=2) + 
  facet_grid(server~side) + theme(legend.position="none")
```